The Sloan Digital Sky Survey Spectroscopic Reduction Pipeline

### 2006-03-31 doc/paper/paper2d.tex
Make it tex.
Added disp figure caption

A plate is mounted in the telescope through a efficient procedure described
in Gunn et al. 2006.  The switch from the imaging camera to a spectroscopic
cartridge takes an average of xx minutes, while a switch from one cartridge
to another takes merely x minutes.  Once the cartridge is mounted, the telescope is slewed to the coordinates of the plate center.  
Before acquisition of the guide stars, the 8 flat-field petals are closed and
a sequence of one 10-second flat-field exposure and one 2-second 
Neon-Mercury-Cadmium arc exposure are obtained.  While the arc-lamp frame is
being readout, the flat-field petals are opened, and guide stars are acquired
through the 11 coherent guide bundles (9 are this big and 2 are even bigger??).
With at least 4 guide stars acquired, closed loop guiding is started, and
usually converges to stable guiding within 2 minutes??
The plate scale of the SDSS telescope can be adjusted, and the guiding software
suggest a scale change if one is required.  Once close loop 
guiding has settled, the first science exposure is started.  The typical 
exposure time is 900 seconds, but exposure times from 600 to 1200 second are
not unexpected based on conditions and time constraints.  A second exposure
is started immediately after the first science exposure is complete and readout.A fast linux workstation onsite copies the data from the data acquisition host,
and automatically performs a quick spectral extraction with a subset of the 
steps explained here.  The results are posted to sos.apo.nmsu.edu, and
the SDSS observers can check the S/N obtained in each of the 4 spectroscopic
cameras in real-time.  A plate is considered to have met SDSS survey standards
if all 4 cameras have a cumulative (S/N)^2 greater than 15 at 
fidicual magnitudes of ?? and ??.

@@

The final flat-field spectra are stored as 
detected counts per fiber per pixel with wavelengths of each pixel
given by the associated solution of the arc-lamp frame.  
We model the flat-field spectra as the product:

\begin{equation} 
{\rm Flat Model}_{ij} = A * f_{avg}(\lambda) \, {\rm fflat_i} (\lambda) 
\end{equation}

where $A$ is a overall normalization in counts, 
$f_{avg}$ is the average response per constant logarithmic wavelength
interval, $fflat_i$ is the fiber-to-fiber response as a function of wavelength.
This model of the extracted flat-field spectra provides a transformation to
compare detected counts from one fiber to another.

@@

Sky subtraction is performed on a frame by frame basis using the 
"flattened" wavelength calibrated extractions.  The goal of sky subtraction
is to estimate the expected background counts contributing to each of 
the native 320x2048 pixels in each extracted frame.  A "supersky" model 
at a fiducial airmass of 1 is constructed by performing iterative fits to
the extracted spectra associated with blank sky positions (so-called
"sky" fibers).   The following steps are taken in 3 sequential iterations
in order to produce the sky-subtracted spectra:

@@

The set of all blue and red exposures for each fiber are
combined using an iterative B-spline fit to reject outlying flux pixels
and remaining cosmic rays.
The splines are evaluated on a wavelength grid that
is uniform is log-wavelengths, with the scale always chosen
as $10^{-4}$ in log-wavelength.  The wavelength ranges are
slightly different for each plate, chosen to span the input data.

@@

The remaining spectro-photometric corrections are functions across
the entire plate that are designed to remove large-spatial scale centering
and guiding errors.
Some of the terms are achromatic with position on the focal plane
(X=XFOCAL and Y=YFOCAL).  There are also chromatic terms that scale
as $ L \equiv 1 - (5100\AA / \lambda)^2 $, since that function gives
an equal effect between $3900$ and $5070\AA$ as between $5070$
and $9000\AA$.
There are also magnitude offsets as a function fo spectrograph ID,
and a chromatic offset as a function of spectrograph ID.
$$ F' = F  \times \left[ a_0 ({\rm SPECID~EQ~1}) + a_1 ({\rm SPECID~EQ~2}) \right]
  \times \exp \left[ a_2 x + a_3 y + a_4 xy + a_5 x^2 + a_6 y^2 + a_7 xL + a_8 yL
  + a_9 L ({\rm SPECID~EQ~1}) + a_10 L ({\rm SPECID~EQ~2})
  + a_11 L^2 ({\rm SPECID~EQ~1}) + a_12 L^2 ({\rm SPECID~EQ~2}) \right] $$
$$ X \equiv {\rm XFOCAL} $$
$$ Y \equiv {\rm YFOCAL} $$
$$ L \equiv 1 - (5100\AA / \lambda)^2 $$

@@

The IAU standard for conversion from air to vacuum wavelengths
is given in Morton (1991, ApJS, 77, 119). For vacuum wavelengths ($\lambda$)
in Angstroms, convert to air wavelength ($\lambda_{\rm air}$) via:
$$\lambda_{\rm air} = \lambda / (1 + 2.735182E-4 + 131.4182 / \lambda^2 + 2.76249E8 / \lambda^4) $$

@@

%------------------------------------------------------------------------------

>>>>

A plate is mounted in the telescope through a efficient procedure described
in Gunn et al. 2006.  The switch from the imaging camera to a spectroscopic
cartridge takes an average of xx minutes, while a switch from one cartridge
to another takes merely x minutes.  Once the cartridge is mounted, the telescope is slewed to the coordinates of the plate center.  
Before acquisition of the guide stars, the 8 flat-field petals are closed and
a sequence of one 10-second flat-field exposure and one 2-second 
Neon-Mercury-Cadmium arc exposure are obtained.  While the arc-lamp frame is
being readout, the flat-field petals are opened, and guide stars are acquired
through the 11 coherent guide bundles (9 are this big and 2 are even bigger??).
With at least 4 guide stars acquired, closed loop guiding is started, and
usually converges to stable guiding within 2 minutes??
The plate scale of the SDSS telescope can be adjusted, and the guiding software
suggest a scale change if one is required.  Once close loop 
guiding has settled, the first science exposure is started.  The typical 
exposure time is 900 seconds, but exposure times from 600 to 1200 second are
not unexpected based on conditions and time constraints.  A second exposure
is started immediately after the first science exposure is complete and readout.A fast linux workstation onsite copies the data from the data acquisition host,
and automatically performs a quick spectral extraction with a subset of the 
steps explained here.  The results are posted to sos.apo.nmsu.edu, and
the SDSS observers can check the S/N obtained in each of the 4 spectroscopic
cameras in real-time.  A plate is considered to have met SDSS survey standards
if all 4 cameras have a cumulative (S/N)$^2$ greater than 15 at 
fidicual magnitudes of ?? and ??.

@@

The final flat-field spectra are stored as 
detected counts per fiber per pixel with wavelengths of each pixel
given by the solution of the matched arc-lamp frame.  
We model the flat-field spectra as the product:

\begin{equation} 
{\rm Model}_{ij} = A \times {\rm SF}(\lambda) \times {\rm (fiber~flat)_i} (\lambda) 
\end{equation}

where $A$ is a overall normalization in counts, 
$SF$ is the average wavelength dependent response called the ``superflat'', and
${\rm fiber~flat_i}$ is the fiber-to-fiber response as a function of wavelength.
This model of the extracted flat-field spectra provides a transformation to
compare detected counts from one fiber to another.

@@

Sky subtraction is performed on a frame by frame basis using the 
``flattened'' wavelength calibrated extractions.  The goal of sky subtraction
is to estimate the expected background counts contributing to each of 
the native 320x2048 pixels in each extracted frame.  A ``supersky'' model 
at a fiducial airmass of 1 is constructed by performing iterative fits to
the extracted spectra associated with blank sky positions (so-called
``sky'' fibers).   The following steps are taken in 3 sequential iterations
in order to produce the sky-subtracted spectra:

@@

The set of all blue and red exposures for each fiber are
combined using an iterative B-spline fit to reject outlying flux pixels
and remaining cosmic rays.
The splines are evaluated on a wavelength grid that
is uniform is log-wavelengths, with the scale always chosen
as $10^{-4}$ in log$_{10}$-wavelength (pixels are constant in 
velocity width with $\Delta v = 69.02$ km/s).  The wavelength ranges are
slightly different for each plate, chosen to span the input data.

@@

The remaining spectro-photometric corrections are functions across
the entire plate that are designed to remove large-spatial scale centering
and guiding errors.
Some of the terms are achromatic with position on the focal plane
(X=XFOCAL and Y=YFOCAL).  There are also chromatic terms that scale
as $ L \equiv 1 - (5100\AA / \lambda)^2 $, since that function gives
an equal effect between $3900$ and $5070\AA$ as between $5070$
and $9000\AA$.
There are also magnitude offsets as a function of spectrograph ID,
and a chromatic offset as a function of spectrograph ID.
$$ F' =  F  \times \left[ a_0 {\rm S_1} + a_1 {\rm S_2} \right]
 \times \exp \left[ a_2 x + a_3 y + a_4 xy + a_5 x^2 + a_6 y^2  + a_7 xL + a_8 yL \right] $$
$$  \times \exp \left[ a_9 L \, {\rm S_1} + a_{10} L \, {\rm S_2}
  + a_{11} L^2 \, {\rm S_1} + a_{12} L^2 \, {\rm S_2} \right]  $$
$$ S_1 \equiv {\rm SPECID~EQ~1} $$
$$ S_2 \equiv {\rm SPECID~EQ~2} $$
$$ X \equiv {\rm XFOCAL} $$
$$ Y \equiv {\rm YFOCAL} $$
$$ L \equiv 1 - (5100\AA / \lambda)^2 $$

@@

The IAU standard for conversion from air to vacuum wavelengths
is given in Morton (1991, ApJS, 77, 119). For vacuum wavelengths ($\lambda$)
in Angstroms, convert to air wavelength ($\lambda_{\rm air}$) via:
$$\lambda_{\rm air} = \lambda / (1 + 0.0002735182 + 131.4182 / \lambda^2 + 2.76249E8 / \lambda^4) $$

@@

%------------------------------------------------------------------------------
%% Figures

\begin{figure}[tb]
\figurenum{7}
\epsscale{0.9}
\plotone{plot_disp.ps}
\figcaption{Measurements of spectral resolution in the four SDSS 
spectrograph cameras.  For each camera, we should the median
FWHM arc-line widths as a function of wavelength as 
solid lines for cameras b1 (cyan) and r1 (orange),
dashed lines for cameras b2 (blue) and r2 (red).
The range of resolution is shown as error bars representing
the 10\% to 90\% ranges in each camera over all fibers and
all plates through DR5.
\label{fig7}
}
\end{figure}
